---
output:
  github_document: default
  html_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

comtradr
=======

[![Travis-CI Build Status](https://travis-ci.org/ChrisMuir/comtradr.svg?branch=master)](https://travis-ci.org/ChrisMuir/comtradr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ChrisMuir/comtradr?branch=master&svg=true)](https://ci.appveyor.com/project/ChrisMuir/comtradr)
[![codecov](https://codecov.io/github/ChrisMuir/comtradr/branch/master/graphs/badge.svg)](https://codecov.io/github/ChrisMuir/comtradr)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/comtradr)](http://cran.r-project.org/package=comtradr)

R package for interacting with the [UN Comtrade Database](https://comtrade.un.org/data/) public API. UN Comtrade provides historical data on the weights and value of 
specific goods shipped between countries, more info can be found [here](https://comtrade.un.org/). Full API documentation can be found 
[here](https://comtrade.un.org/data/doc/api/).

This package was inspired by the [R tutorial](https://comtrade.un.org/data/Doc/api/ex/r) posted by Comtrade, and is built using
[httr](https://CRAN.R-project.org/package=httr) and [jsonlite](https://CRAN.R-project.org/package=jsonlite).

I've also built a Shiny app for visualizing comtrade shipping data, that's powered by this package. The app can be viewed [here](https://chrismuir.shinyapps.io/comtrade_plot_shinyapp/).

Please [report](https://github.com/ChrisMuir/comtradr/issues) issues, comments, or feature requests.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

**NOTE**: As of 2017-06-27, the curl errors due to an inadility to validate the SSL certificate of the API site are no longer appearing, I believe that UN Comtrade fixed their SSL cert issue. As of 2017-07-03, the package is back on CRAN.

## Installation

Install from CRAN:
```{r eval=FALSE}
install.packages("comtradr")
```

Or install from this repo:
```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("ChrisMuir/comtradr")
```

## Example Usage

**Example 1**: Return all exports from China to South Korea, United States and Mexico, for all years

```{r}
library(comtradr)

# First, read in the country code lookup table as a dataframe from Comtrade. 
# This will be used as a parameter in the API calls.
countrydf <- ct_countries_table()

# This object can also be used to look up the exact spellings of countries 
# prior to making API calls.
ct_country_lookup("korea", "reporter", countrydf)

# Since we want South Korea, we'll use "Rep. of Korea" within the API query.
example1 <- ct_search(reporters = "China", 
                      partners = c("Rep. of Korea", "USA", "Mexico"), 
                      countrytable = countrydf, 
                      tradedirection = "exports")

# Inspect the return data
str(example1$data)
```

**Example 2**: Return all exports related to shrimp from Thailand to all other countries, for years 2007 thru 2011

```{r}
library(comtradr)

# First, read in the commodity code lookup table as a dataframe from Comtrade.
commoditydf <- ct_commodities_table("HS")

# Then search for shrimp.
shrimp_codes <- ct_commodity_lookup("shrimp", commoditydf, return_code = TRUE, return_char = TRUE)

# API query.
example2 <- ct_search(reporters = "Thailand", 
                      partners = "All", 
                      countrytable = countrydf, 
                      tradedirection = "exports", 
                      startdate = "2007-01-01", 
                      enddate = "2011-01-01", 
                      commodcodes = shrimp_codes)

# Inspect the output
str(example2$data)
```
