---
title: "comtradr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comtradr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comtradr)
```


## data availablility

See here for an overview of available commodity classifications:  https://unstats.un.org/wiki/display/comtrade/New+Comtrade+FAQ+for+Advanced+Users?preview=/135004494/135004492/Picture1.png



## Package information

API wrapper for the [UN Comtrade Database](https://comtradeplus.un.org/). UN Comtrade provides historical data on the weights and value of specific goods shipped between countries, more info can be found [here](https://unstats.un.org/wiki/display/comtrade). Full API documentation can be found [here](https://comtradedeveloper.un.org/).

## Install and load comtradr

Install the development version from GitHub:
```{r eval = FALSE}
devtools::install_github("ChrisMuir/comtradr")
```

Load comtradr
```{r}
library(comtradr)
```

<!-- ## Making API calls -->
<!-- Lets say we want to get data on all imports into the United States from Germany, France, Japan, and Mexico, for all years. -->
<!-- ```{r, echo = FALSE} -->
<!-- v_data_1 <- system.file("extdata", "vignette_data_1.rda", package = "comtradr") -->
<!-- if (!file.exists(v_data_1)) { -->
<!--   stop("internal vignette data set '~/extdata/vignette_data_1.rda' not found", call. = FALSE) -->
<!-- } -->
<!-- load(v_data_1) -->
<!-- ``` -->

<!-- ```{r, eval = FALSE} -->
<!-- q <- ct_get_data(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Japan", "Mexico"),  -->
<!--                trade_direction = "imports") -->
<!-- ``` -->
<!-- API calls return a tidy data frame. -->
<!-- ```{r} -->
<!-- str(q) -->
<!-- ``` -->

<!-- Here are a few more examples to show the different parameter options: -->

<!-- Limit the search range to shipments between 2010 and 2014. -->
<!-- ```{r, eval = FALSE} -->
<!-- q <- ct_search(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Japan", "Mexico"),  -->
<!--                trade_direction = "imports",  -->
<!--                start_date = 2010,  -->
<!--                end_date = 2014) -->
<!-- ``` -->

<!-- By default, the return data is in yearly amounts. We can pass `"monthly"` to arg `freq` to return data in monthly amounts, however the API limits each "monthly" query to a single year. -->
<!-- ```{r, eval = FALSE} -->
<!-- # Get all monthly data for a single year (API max of 12 months per call). -->
<!-- q <- ct_search(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Japan", "Mexico"),  -->
<!--                trade_direction = "imports",  -->
<!--                start_date = 2012,  -->
<!--                end_date = 2012,  -->
<!--                freq = "monthly") -->

<!-- # Get monthly data for a specific span of months (API max of five months per call). -->
<!-- q <- ct_search(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Japan", "Mexico"),  -->
<!--                trade_direction = "imports",  -->
<!--                start_date = "2012-03",  -->
<!--                end_date = "2012-07",  -->
<!--                freq = "monthly") -->
<!-- ``` -->

<!-- Countries passed to parameters `reporters` and `partners` must be spelled as they appear in the Comtrade country reference table. Function `ct_country_lookup` allows us to query the country reference table. -->
<!-- ```{r} -->
<!-- ct_country_lookup("korea", "reporter") -->
<!-- ct_country_lookup("bolivia", "partner") -->
<!-- ``` -->
<!-- ```{r, eval = FALSE} -->
<!-- q <- ct_search(reporters = "Rep. of Korea",  -->
<!--                partners = "Bolivia (Plurinational State of)",  -->
<!--                trade_direction = "all") -->
<!-- ``` -->

<!-- Search trade related to specific commodities (say, tomatoes). We can query the Comtrade commodity reference table to see all of the different commodity descriptions available for tomatoes. -->
<!-- ```{r} -->
<!-- ct_commodity_lookup("tomato") -->
<!-- ``` -->
<!-- If we want to search for shipment data on all of the commodity descriptions listed, then we can simply adjust the parameters for `ct_commodity_lookup` so that it will return only the codes, which can then be passed along to `ct_search`. -->
<!-- ```{r, eval = FALSE} -->
<!-- tomato_codes <- ct_commodity_lookup("tomato",  -->
<!--                                     return_code = TRUE,  -->
<!--                                     return_char = TRUE) -->

<!-- q <- ct_search(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Mexico"),  -->
<!--                trade_direction = "all",  -->
<!--                commod_codes = tomato_codes) -->
<!-- ``` -->
<!-- On the other hand, if we wanted to exclude juices and sauces from our search, we can pass a vector of the relevant codes to the API call. -->
<!-- ```{r, eval = FALSE} -->
<!-- q <- ct_search(reporters = "USA",  -->
<!--                partners = c("Germany", "France", "Mexico"),  -->
<!--                trade_direction = "all",  -->
<!--                commod_codes = c("0702", "070200", "2002", "200210", "200290")) -->
<!-- ``` -->

<!-- ## API search metadata -->

<!-- In addition to the trade data, each API return object contains metadata as attributes. -->
<!-- ```{r} -->
<!-- # The url of the API call. -->
<!-- attributes(q)$url -->
<!-- # The date-time of the API call. -->
<!-- attributes(q)$time_stamp -->

<!-- # The total duration of the API call, in seconds. -->
<!-- attributes(q)$req_duration -->
<!-- ``` -->

<!-- ## More on the lookup functions -->

<!-- Functions `ct_country_lookup` and `ct_commodity_lookup` are both able to take multiple search terms as input. -->
<!-- ```{r} -->
<!-- ct_country_lookup(c("Belgium", "vietnam", "brazil"), "reporter") -->

<!-- ct_commodity_lookup(c("tomato", "trout"), return_char = TRUE) -->
<!-- ``` -->


<!-- `ct_commodity_lookup` can return a vector (as seen above) or a named list, using parameter `return_char` -->
<!-- ```{r} -->
<!-- ct_commodity_lookup(c("tomato", "trout"), return_char = FALSE) -->
<!-- ``` -->


<!-- For `ct_commodity_lookup`, if any of the input search terms return zero results and parameter `verbose` is set to `TRUE`, a warning will be printed to console (set `verbose` to `FALSE` to turn off this feature). -->
<!-- ```{r} -->
<!-- ct_commodity_lookup(c("tomato", "sldfkjkfdsklsd"), verbose = TRUE) -->
<!-- ``` -->

<!-- ## API rate limits -->

<!-- The Comtrade API imposes rate limits on both guest users and premium users. `comtradr` features automated throttling of API calls to ensure the user stays within the limits defined by Comtrade. Below is a breakdown of those limits, API docs on these details can be found [here](https://comtrade.un.org/data/doc/api/#Authentication). -->

<!-- * Without user token: 1 request per second, 100 requests per hour. -->
<!-- * With valid user token: 1 request per second, 10,000 requests per hour. -->

<!-- In addition to these rate limits, the API imposes some limits on parameter combinations. -->

<!-- * Between args `reporters`, `partners`, and the query date range, only one of these three may use the catch-all input "All". -->
<!-- * For the same group of three (`reporters`, `partners`, date range), if the input is not "All", then the maximum number of input values for each is five. For date range, if not using "All", then the `start_date` and `end_date` must not span more than five months or five years. There is one exception to this rule, if arg `freq` is "monthly", then a single year can be passed to `start_date` and `end_date` and the API will return all of the monthly data for that year. -->
<!-- * For arg `commod_codes`, if not using input "All", then the maximum number of input values is 20 (although "All" is always a valid input). -->

<!-- Additionally, the default maximum number of returned records from a single query is 100,000. If you have a valid registered token, and have a higher per-query limit, you can pass in a max int to input param `max_rec`. -->

<!-- `comtradr` features a few functions for working with the API rate limits and tokens. -->

<!-- * `ct_register_token()` allows the user to set an API token within the package environment.  -->
<!-- * `ct_get_remaining_hourly_queries()` will return the number of remaining queries for the current hour. -->
<!-- * `ct_get_reset_time()` will return the date/time in which the current hourly time limit will reset, as a `POSIXct` object. -->

<!-- ## Package Data -->

<!-- `comtradr` ships with a few different package data objects, and functions for interacting with and using the package data. -->

<!-- **Country/Commodity Reference Tables** -->

<!-- As explained previously, making API calls with `comtradr` often requires the user to query the country reference table and/or the commodity reference table (this is done using functions `ct_country_lookup` and `ct_commodity_lookup`). Both of these reference tables are generated by the UN Comtrade, and are updated roughly once a year. Since they're updated infrequently, both tables are saved as cached data objects within the `comtradr` package, and are referenced by the package functions when needed. -->

<!-- `comtradr` features a function, `ct_update_databases`, for checking the Comtrade website for updates to either reference table. If updates are found, the function will download the updated table, save it to the package directory, and make it available during the current R session. It will also print a message indicating whether updates were found, like so: -->
<!-- ```{r, eval = FALSE} -->
<!-- ct_update_databases() -->
<!-- #> All DB's are up to date, no action required -->
<!-- ``` -->
<!-- If any updates are found, the message will state which reference table(s) were updated. -->

<!-- The user may force download of both reference tables (regardless of whether updates exist) by using arg `force = TRUE` within function `ct_update_databases`. This is useful in the event that either reference table file is deleted or removed from the package directory. If this is the case, and either reference table file is not found upon package load, then any subsequent `comtradr` functions that require the use of a reference table will result in an error, with the error message prompting the user to run `ct_update_databases(force = TRUE)`. -->

<!-- Additionally, the Comtrade API features a number of different commodity reference tables, based on different trade data classification schemes (for more details, see [this](https://comtrade.un.org/data/doc/api/#DataAvailabilityRequests) page from the API docs). `comtradr` ships with the commodity table for the "Harmonized System", or "HS", scheme. The user may download any of the available commodity tables by specifying arg `commodity_type` within function `ct_update_databases` (e.g., `ct_update_databases(commodity_type = "SITC")` will download the commodity table that follows the "Standard International Trade Classification" scheme). Doing this will replace the commodity table on file with the one specified. To see the classification scheme of the commodity table currently on file, use `ct_commodity_db_type`. -->
<!-- ```{r} -->
<!-- ct_commodity_db_type() -->
<!-- ``` -->

<!-- **"Polished" Column Headers** -->

<!-- `ct_pretty_cols` is a named vector of column header values that provide the option of using column headers that are more polished and human-friendly than those returned by the API function `ct_search`. The polished column headers may be useful when plotting the Comtrade data, or for use in publication tables. The data can be accessed directly by using `data("ct_pretty_cols")`, but there is also a function for applying the polished headers to `comtradr` data frames, `ct_use_pretty_cols`. Below is a quick demonstration. -->

<!-- ```{r} -->
<!-- # Column headers returned from function ct_search -->
<!-- colnames(q) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # Apply polished column headers -->
<!-- q <- ct_use_pretty_cols(q) -->

<!-- # Print new column headers. -->
<!-- colnames(q) -->
<!-- ``` -->

<!-- ## Visualize -->

<!-- Once the data is collected, we can use it to create some basic visualizations. -->

<!-- **Plot 1**: Plot total value (USD) of Chinese exports to Mexico, South Korea and the United States, by year. -->

<!-- ```{r, echo = FALSE} -->
<!-- v_data_2 <- system.file("extdata", "vignette_data_2.rda", package = "comtradr") -->
<!-- if (!file.exists(v_data_2)) { -->
<!--   stop("internal vignette data set '~/extdata/vignette_data_2.rda' not found", call. = FALSE) -->
<!-- } -->
<!-- load(v_data_2) -->
<!-- ``` -->
<!-- ```{r, eval = FALSE} -->
<!-- # Comtrade api query. -->
<!-- df <- ct_search(reporters = "China",  -->
<!--                 partners = c("Rep. of Korea", "USA", "Mexico"),  -->
<!--                 trade_direction = "exports") -->
<!-- ``` -->

<!-- ```{r, warning = FALSE, message = FALSE} -->
<!-- library(ggplot2) -->

<!-- # Apply polished col headers. -->
<!-- df <- ct_use_pretty_cols(df) -->

<!-- # Create plot. -->
<!-- ggplot(df, aes(Year, `Trade Value usd`, color = factor(`Partner Country`),  -->
<!--                shape = factor(`Partner Country`))) + -->
<!--   geom_point(size = 2) + -->
<!--   geom_line(size = 1) + -->
<!--   scale_x_continuous(limits = c(min(df$Year), max(df$Year)),  -->
<!--                      breaks = seq.int(min(df$Year), max(df$Year), 2)) + -->
<!--   scale_color_manual(values = c("orange", "blue", "red"),  -->
<!--                      name = "Destination\nCountry") + -->
<!--   scale_shape_discrete(name = "Destination\nCountry") + -->
<!--   labs(title = "Total Value (USD) of Chinese Exports, by Year") + -->
<!--   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) -->
<!-- ``` -->

<!-- **Plot 2**: Plot the top eight destination countries/areas of Thai shrimp exports, by weight (KG), for 2007 - 2011. -->

<!-- ```{r, echo = FALSE} -->
<!-- v_data_3 <- system.file("extdata", "vignette_data_3.rda", package = "comtradr") -->
<!-- if (!file.exists(v_data_3)) { -->
<!--   stop("internal vignette data set '~/extdata/vignette_data_3.rda' not found", call. = FALSE) -->
<!-- } -->
<!-- load(v_data_3) -->
<!-- ``` -->
<!-- ```{r, eval = FALSE} -->
<!-- # First, collect commodity codes related to shrimp. -->
<!-- shrimp_codes <- ct_commodity_lookup("shrimp",  -->
<!--                                     return_code = TRUE,  -->
<!--                                     return_char = TRUE) -->

<!-- # Comtrade api query. -->
<!-- df <- ct_search(reporters = "Thailand",  -->
<!--                 partners = "All",  -->
<!--                 trade_direction = "exports",  -->
<!--                 start_date = 2007,  -->
<!--                 end_date = 2011,  -->
<!--                 commod_codes = shrimp_codes) -->
<!-- ``` -->

<!-- ```{r, warning = FALSE, message = FALSE} -->
<!-- library(ggplot2) -->
<!-- library(dplyr) -->

<!-- # Apply polished col headers. -->
<!-- df <- ct_use_pretty_cols(df) -->

<!-- # Create country specific "total weight per year" dataframe for plotting. -->
<!-- plotdf <- df %>%  -->
<!--   group_by_(.dots = c("`Partner Country`", "Year")) %>%  -->
<!--   summarise(kg = as.numeric(sum(`Net Weight kg`, na.rm = TRUE))) %>%  -->
<!--   as_data_frame() -->

<!-- # Get vector of the top 8 destination countries/areas by total weight shipped  -->
<!-- # across all years, then subset plotdf to only include observations related  -->
<!-- # to those countries/areas. -->
<!-- top8 <- plotdf %>%  -->
<!--   group_by(`Partner Country`) %>%  -->
<!--   summarise(kg = as.numeric(sum(kg, na.rm = TRUE))) %>%  -->
<!--   top_n(8, kg) %>% -->
<!--   arrange(desc(kg)) %>%  -->
<!--   .[["Partner Country"]] -->
<!-- plotdf <- plotdf %>% filter(`Partner Country` %in% top8) -->

<!-- # Create plots (y-axis is NOT fixed across panels, this will allow us to ID  -->
<!-- # trends over time within each country/area individually). -->
<!-- qplot(Year, kg, data = plotdf) +  -->
<!--   geom_line(data = plotdf[plotdf$`Partner Country` %in% names(which(table(plotdf$`Partner Country`) > 1)), ]) +  -->
<!--   xlim(min(plotdf$Year), max(plotdf$Year)) +  -->
<!--   labs(title = "Weight (KG) of Thai Shrimp Exports, by Destination Area, 2007 - 2011") +  -->
<!--   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),  -->
<!--         axis.text = element_text(size = 7)) +  -->
<!--   facet_wrap(~factor(`Partner Country`, levels = top8), scales = "free", nrow = 2, ncol = 4) -->
<!-- ``` -->
