---
title: "The estimated trade matrix"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{trade_matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comtradr)
```

## What is the estimated trade matrix?

`ct_get_data()` gives you exactly what countries have *reported* to the UN
Comtrade Database, and nothing else. The main database
[contains no estimates](https://uncomtrade.org/docs/read-me-first-disclaimer/).
That is usually what you want — but it means a country that filed nothing
simply is not there, and a world total built from reported data is silently
short by however much that country trades.

The **Trade Matrix** endpoint is a separate analytical product that fills
those gaps. It is maintained by the UN Statistics Division, which
complements reported figures with estimates so that the reporter × partner
matrix covers world trade completely. `ct_get_trade_matrix()` queries it.

How much difference does that make? For food and live animals (SITC section
`"0"`) in 2022:

```{r, echo = FALSE}
v_data_tm <- system.file(
  "extdata", "vignette_data_trade_matrix.rda",
  package = "comtradr"
)
if (!file.exists(v_data_tm)) {
  stop(
    "internal vignette data set ",
    "'~/extdata/vignette_data_trade_matrix.rda' not found",
    call. = FALSE
  )
}
load(v_data_tm)
```

```{r}
trade_matrix_coverage
```

43 of those 209 reporters have no reported figure anywhere in this matrix —
every one of their flows is a UN estimate. They account for only about 3% of
world section-`"0"` exports by value, but that is not the point. The point is
*which* countries they are (the ten largest are shown here):

```{r}
trade_matrix_nonreporters
```

If your analysis involves Russia, Belarus or Honduras in 2022, `ct_get_data()`
will not tell you they are missing. It will just return a smaller matrix. That
is the case for reaching for this endpoint.

One caution, which the next-but-one section explains: "no reported figure in
this matrix" is not quite the same as "filed nothing". Iraq also appears on
the full list, yet it *did* file 2022 data with Comtrade — the matrix
estimated its food exports regardless. Read the list as *countries whose
figures here are entirely estimated*, and check individual cases against
`ct_get_data()` when the distinction matters.

The rest of this vignette covers the things about it that are easy to get
wrong.

## Authentication

Just like `ct_get_data()`, this function needs a Comtrade API token. If you
have not set one up yet, head over to the main
[`comtradr` vignette](comtradr.html), which walks through obtaining and
storing your key.

## A basic call

Let's build the world **export** matrix for food and live animals in 2022.
Leaving `reporter` and `partner` at their default `"everything"` returns the
full matrix, estimates included.

```{r, eval = FALSE}
matrix_0 <- ct_get_trade_matrix(
  commodity_code = "0",
  flow_direction = "export",
  start_date = 2022,
  end_date = 2022
)
```

To keep this vignette fast and network-free, we do not call the API when it
is built. The objects shown throughout were computed from that same query —
`data-raw/DATASET.R` runs it once with `include_world = TRUE`, so that the
World rows are available for the next section, and derives everything else
from the bilateral rows.

`trade_matrix_sample` is a small illustrative slice — Montenegro's exports in
2022, a single reporter whose partner flows mix reported values with UN
estimates. It keeps only a handful of display columns:

```{r}
str(trade_matrix_sample)
```

```{r}
head(trade_matrix_sample)
```

## The World rows, and why they will ruin your totals

This is the one that bites hardest, so it comes first.

The endpoint returns aggregate **World** rows — reporter code `0`, partner
code `0`, ISO `W00` — interleaved with the bilateral flows. They are the row
margins, the column margins and the grand total of the matrix. They are
labeled `World` in `reporter_desc` / `partner_desc`, but no *flag* column
marks them as aggregates, so nothing stops you summing straight over them.

Here is what the full 2022 section-`"0"` export response actually contains:

```{r}
trade_matrix_blocks
```

Four blocks, each summing to the *same* world total. Add up
`primary_value` over the response as it arrives from the API and you get
four times the right answer.

`ct_get_trade_matrix()` therefore drops these rows by default. If you want
the margins or the grand total, ask for them explicitly. (One exception: if
you request `partner = "World"`, that margin is the thing you asked for and
is kept — the reporter margin and grand total are still dropped, so the
result is still safe to sum.)

```{r, eval = FALSE}
# bilateral flows only -- safe to aggregate (the default)
bilateral <- ct_get_trade_matrix(
  commodity_code = "0", flow_direction = "export",
  start_date = 2022, end_date = 2022
)

# margins and grand total included -- do NOT sum across the whole frame
with_margins <- ct_get_trade_matrix(
  commodity_code = "0", flow_direction = "export",
  start_date = 2022, end_date = 2022,
  include_world = TRUE
)
```

**Do not try to filter these rows using `is_aggregate`.** The column exists
in the response, but this endpoint never populates it: it is `FALSE` on every
row, World rows included. Reporter or partner code `0` is the only reliable
marker.

## What `is_reported` actually means

The obvious reading of `is_reported` is "TRUE if the country reported this,
FALSE if the UN estimated it because the country did not report". That
reading is wrong, and it will lead you astray.

`is_reported` is a **per-cell provenance flag**:

- `TRUE` — the value is the reporting country's own figure.
- `FALSE` — the cell was produced or adjusted by the UNSD estimation
  pipeline: extrapolated from a nearby reported year, derived by mirror
  inversion from partner data, manually adjusted for under-reported or
  confidential trade, or produced by redistributing non-specified partners.

Three consequences worth internalizing:

**A country that reports fully still has `FALSE` cells.** Germany reported
2022 in full, yet 77 of its 233 partner cells are flagged `FALSE` — and 63 of
those agree with the reported figure to within rounding (a median difference
of about \$16,000 on values in the millions). Only 14 are revised materially,
all upward. The flag marks passage through the estimation pipeline, not
necessarily a changed number.

**Aggregates inherit the flag.** A `TOTAL` row is `FALSE` if *any* underlying
section was estimated, even where every other section was reported as filed.

**So the reported/estimated split is not "extra coverage".** Over the
bilateral rows of our 2022 section-`"0"` matrix:

```{r}
trade_matrix_totals
```

That `FALSE` bucket — about 7% of the total — is not "trade the UN estimated
for countries that had not reported". It is every cell that passed through
the pipeline for any reason. Do not present it as the coverage gap; the
coverage gap is the 43 countries shown at the top of this vignette.

To find those countries, group by reporter and ask whether *any* row is
reported, rather than reading the flag row by row:

```{r, eval = FALSE}
library(dplyr)

non_reporters <- matrix_0 |>
  group_by(reporter_iso, reporter_desc) |>
  summarise(any_reported = any(is_reported), .groups = "drop") |>
  filter(!any_reported)
```

## Commodities

Commodities are classified by SITC, and `classification_code` is `"SS"` on
every row — the "combined" SITC list that spans revisions. (The `TM` you see
in the request URL is the endpoint's dataset selector, not a classification.)
The estimation itself is done on SITC Rev.3.

`commodity_code` accepts:

- `"TOTAL"` — all commodities combined.
- One-, two- and three-digit SITC codes, e.g. `"0"`, `"01"`, `"011"`.
- The five four-/five-digit codes the UN estimates in addition: `"7812"`,
  `"7841"`, `"7851"`, `"7852"`, `"78531"`.
- The level selectors `"ag1"` through `"ag5"`, which return every code with
  that many digits. `everything` is a synonym for `"ag1"`.
- `"all_levels"`, which returns the entire hierarchy at once.

The one-digit sections are:

| Code | SITC section |
|:----:|:-------------|
| 0 | Food and live animals |
| 1 | Beverages and tobacco |
| 2 | Crude materials, inedible, except fuels |
| 3 | Mineral fuels, lubricants and related materials |
| 4 | Animal and vegetable oils, fats and waxes |
| 5 | Chemicals and related products |
| 6 | Manufactured goods classified chiefly by material |
| 7 | Machinery and transport equipment |
| 8 | Miscellaneous manufactured articles |
| 9 | Commodities not classified elsewhere |

A second double-counting trap lurks here. `"all_levels"` returns `TOTAL`, the
sections, *and* the two- and three-digit codes nested inside them — so
summing across that result over-counts just as the World rows do. The
function warns you when you ask for it. If you want "all ten sections", use
`everything` or `"ag1"`, which return one level only.

## Values, dates and flows

**Only `primary_value` is populated**, in current US dollars, following the
usual Comtrade convention: CIF-type for imports, FOB-type for exports. The
`cifvalue`, `fobvalue`, `qty` and `net_wgt` columns are always `NA` here.
That is by design — the estimation is applied to trade value only, and
quantity information is stripped from the source data.

**Annual goods only.** `start_date` and `end_date` must be plain years
(`yyyy`), as a number or a string; `2022` and `"2022"` both work. Passing
`"2022-01"` is rejected rather than quietly interpreted as a year. The API
accepts at most 12 periods per query.

**Imports and exports only.** The trade matrix carries no re-exports or other
flows, so `flow_direction` accepts `"import"`, `"export"` or `everything`.
Other values from `ct_get_ref_table('flow_direction')` are rejected up front,
rather than returning a confusingly empty result.

## Caveats worth carrying into your analysis

These come from the UN's own
[Note on the Trade Estimation](https://uncomtrade.org/wp-content/uploads/2026/04/Note-on-Trade-Estimation-26-July-2010-2026-Edit-Public.pdf):

- Mirror-derived estimates "may not fully reflect the data collected and
  compiled by the countries", because trade flows reported by partners do not
  match those reported by the compiling country.
- The manual adjustment of under-reported and confidential trade "is based on
  the best judgment of the staff" at UNSD's Business and Trade Statistics
  Section.
- Aggregate totals from this matrix usually run slightly higher than the
  reported series.
- Because the main database holds no estimates, `ct_get_data()` and
  `ct_get_trade_matrix()` are **not interchangeable**. Figures from the trade
  matrix should not be presented as reported statistics.

Finally, note that this endpoint is not part of the UN Comtrade public API
documentation, which is why `ct_get_trade_matrix()` is marked experimental.
Its field semantics may change without notice.

When publishing, cite the data source as "UN Comtrade".

## See also

- `ct_get_data()` — the standard endpoint, which returns only reported
  values. Reach for it when you want exactly what countries filed, and for
  `ct_get_trade_matrix()` when a missing country would distort your answer.
- The main [`comtradr` vignette](comtradr.html) for token setup, lookup
  helpers and general usage.
